Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-04-28 manuscript/manuscriptRW.Rmd
added more changes from Haralds feedback

---
title: "Tree regeneration in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Friedrich Bohn,
- Kristin H. Braziunas,
- Josef	Brůna,
- Rico Fischer,
- Samuel M. Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Ján Merganič,
- Katarina Merganičová,
- Tobias Mette,
- Marco Mina,
- Xavier Morin,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994]. Overall, models are needed to
(1) synthesize existing empirical data and explore their relationships,
(2) assess future tree regeneration e.g. in the context of global change
scenarios, and (3) identify the most important processes that are
shaping ecological patterns. Given the current strategies that are used
in models of forest dynamics to represent tree regeneration, model
behavior often is prone to problems, such as very high levels of tree
regeneration that necessitate excess mortality at early stages of tree
life to simulate correct stand structure and composition. Also,
inadequately high species diversity in tree regeneration may be
simulated, which is characteristic of "classical" forest gap models
[@gutiérrez2016], at least as long as the simulation set-up comprises a
multitude of species. Some models use calibration against local
canopy-level data to constrain simulated regeneration levels, which is
likely to hamper the general applicability of these models e.g. under
scenarios of climate change. Furthermore, correctly capturing the
species composition in tree regeneration as a function of the presence
of seed trees in the canopy is often a particular challenge, potentially
leading to unrealistic successional drift in the model, which must be
corrected by factors that are hard to parameterize [e.g., @lischke2006].
A related issue is the excessive reduction of species diversity due to
positive feedback effects, such that eventually just single-species
stands remain. This is sometimes corrected by the incorporation of a low
level of seed influx of all species at all times [@schumacher2006] or by
restricting the number of seeds per species in the seedbank
[@lischke2006]. However, simulated species composition is usually
exceedingly sensitive to assumptions about seed availability, whereas
the parameters of such functions are poorly constrained by field data.
Lastly, there is often a problem with insufficient observational
constraints on parameter values for models that start from very small
tree sizes (e.g., 10 cm of height or even seeds) and track tree
development in a process-oriented manner by considering a multitude of
ecological influences, rather than emphasizing tree regeneration into a
larger size class [e.g., @wehrli2006].

@@

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2022]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated
explicitly in most models, such as deer browsing --- we do not aim for a
detailed statistical evaluation of each model. Instead, we aim to
evaluate the general regeneration patterns and magnitudes simulated by
the models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

@@

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = "Models included and their characteristics.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "8em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

@@

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2022 to provide regeneration
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Regeneration rates per site,
sample and decade ranged from zero to 1246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this unique dataset and the detailed regeneration
information, cf. @käber2022.

@@

Mortality in tree regeneration was assessed based on the ratio of
regeneration between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
to estimate whether the ratio of regeneration between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate.
The Reineke self-thinning rule is usually calculated for even-aged,
single species stands and assumes a fixed relationship between the
number of stems and the quadratic mean diameter in fully stocked pure
stands. The value used in our comparisons was 1.77 (i.e., we expect stem
numbers at 7 cm to be 77% higher than at 10 cm), calculated using
Equation \ref{eqReineke}.

@@

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented for one examplary model of each category (overperdicted, intermediate and underpredicted). The full data with the models grouping are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

@@

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. For comparing the magnitude of recruitment across models the same graphs with equal y-axis cf. Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

@@

The EuFoRIa data [@käber2022] are exceptional, particularly with respect
to the number of records (number of sites and repeated measurements),
which is essential for capturing a highly 'noisy' process such as tree
regeneration. The use of this dataset for model benchmarking provided
novel insights on the ability of state-of-the-art models to accurately
simulate regeneration levels, species composition, and mortality in an
early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
Overall, by adopting this approach a much broader understanding resulted
than if we had used these data for model calibration: it is primarily
from the shortcomings of the models that we can gain ecological insights
[cf. @trugman2022].

>>>>

---
title: "Tree regeneration in models of forest dynamics: a key priority for further research"
author:
- Olalla Díaz-Yáñez,
- Yannek Käber,
- Tim Anders,
- Friedrich Bohn,
- Kristin H. Braziunas,
- Josef	Brůna,
- Rico Fischer,
- Samuel M. Fischer,
- Jessica Hetzer,
- Thomas Hickler,
- Christian	Hochauer,
- Manfred J. Lexer,
- Heike	Lischke,
- Mats Mahnken,
- Paola	Mairota,
- Ján Merganič,
- Katarina Merganičová,
- Tobias Mette,
- Marco Mina,
- Xavier Morin,
- Werner Rammer,
- Christopher P.O. Reyer,
- Simon	Scheiter,
- Daniel Scherrer,
- Harald Bugmann
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    number_sections: yes
    toc: no
    extra_dependencies: "subfig"
  word_document:
    reference_docx: template_paper.docx
mainfont: Arial
fontsize: 12pt
linestretch: 1.5
bibliography: Regeneration.bib
link-citations: yes
nocite: |
  @bugmann1996a, @bohn2014, @fabrika2005, @mette2014, @smith2001a, @langan2017, @seidl2012a, @lischke2006, @schumacher2006, @scheiter2009, @scheller2007, @morin2021, @lasch-born2020, @hickler2012, @fischer_lessons_2016, @langan2017, @scheiter2013 
linkcolor: blue
csl: global-change-biology.csl
editor_options:
  markdown:
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{caption}
- \captionsetup[figure]{font=scriptsize}
- \captionsetup[table]{font=scriptsize}
---

@@

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shugart1984] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994]. Overall, models are needed to
(1) synthesize existing empirical data and explore their relationships,
(2) assess future tree regeneration e.g. in the context of global change
scenarios, and (3) identify the most important processes that are
shaping ecological patterns. Given the current strategies that are used
in models of forest dynamics to represent tree regeneration, model
behavior often is prone to problems, such as very high levels of tree
regeneration that necessitate excess mortality at early stages of tree
life to simulate correct stand structure and composition. Also,
inadequately high species diversity in tree regeneration may be
simulated, which is characteristic of "classical" forest gap models
[@gutiérrez2016], at least as long as the simulation set-up comprises a
multitude of species. Some models use calibration against local
canopy-level data to constrain simulated regeneration levels, which is
likely to hamper the general applicability of these models e.g. under
scenarios of climate change. Furthermore, correctly capturing the
species composition in tree regeneration as a function of the presence
of seed trees in the canopy is often a particular challenge, potentially
leading to unrealistic successional drift in the model, which must be
corrected by factors that are hard to parameterize [e.g., @lischke2006].
A related issue is the excessive reduction of species diversity due to
positive feedback effects, such that eventually just single-species
stands remain. This is sometimes corrected by the incorporation of a low
level of seed influx of all species at all times [@schumacher2006] or by
restricting the number of seeds per species in the seedbank
[@lischke2006]. However, simulated species composition is usually
exceedingly sensitive to assumptions about seed availability, whereas
the parameters of such functions are poorly constrained by field data.
Lastly, there is often a problem with insufficient observational
constraints on parameter values for models that start from very small
tree sizes (e.g., 10 cm of height or even seeds) and track tree
development in a process-oriented manner by considering a multitude of
ecological influences, rather than emphasizing tree regeneration into a
larger size class [e.g., @wehrli2006].

@@

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2023]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated
explicitly in most models, such as deer browsing --- we do not aim for a
detailed statistical evaluation of each model. Instead, we aim to
evaluate the general regeneration patterns and magnitudes simulated by
the models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

@@

kable(modelDF,
      format = "latex", booktabs = TRUE,
      caption = "Models included and their characteristics.") |>
        kableExtra::landscape() |>
        kableExtra::kable_styling(full_width = F, latex_options = "scale_down") |>
    # fix width by column
      kableExtra::column_spec(column = 1, width = "5em") |>
      kableExtra::column_spec(column = 2, width = "3em") |>
      kableExtra::column_spec(column = 3, width = "4em") |>
      kableExtra::column_spec(column = 4, width = "5em") |>
      kableExtra::column_spec(column = 5, width = "5em") |>
      kableExtra::column_spec(column = 6, width = "5em") |>
      kableExtra::column_spec(column = 7, width = "8em") |>
      kableExtra::column_spec(column = 8, width = "8em") |>
      kableExtra::column_spec(column = 9, width = "5em") |>
      kableExtra::column_spec(column = 10, width = "5em") |>
      kableExtra::column_spec(column = 11, width = "10em") |>
      kableExtra::add_footnote(c("S-Stand, L-Landscape, G-Global",
                                 "PB-Process based, EM-Empirical",
                                 "C-Cohort, I-Individual",
                                 "Y-Yes, N-No",
                                 "D-Daily, M-Monthly"),

@@

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2023 to provide regeneration
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Regeneration rates per site,
sample and decade ranged from zero to 1246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this unique dataset and the detailed regeneration
information, cf. @käber2023.

@@

Mortality in tree regeneration was assessed based on the ratio of
regeneration between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2005] as a reference
to estimate whether the ratio of regeneration between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate.
The Reineke self-thinning rule is usually calculated for even-aged,
single species stands and assumes a fixed relationship between the
number of stems and the quadratic mean diameter in fully stocked pure
stands. The value used in our comparisons was 1.77 (i.e., we expect stem
numbers at 7 cm to be 77% higher than at 10 cm), calculated using
Equation \ref{eqReineke}.

@@

```{r combinedRichness, echo=FALSE, fig.cap="Mean Shannon index across all samples per site for observed and simulated data. Each plot shows one pattern represented by one exemplary model of each category (overperdicted, intermediate and underpredicted). The full data with the grouping of the models are shown in Figures S3 and S4. n indicates the number of models falling in each group. A) Examples for the three trends across models for species diversity at the stand level. B) Examples for the three trends across models for regeneration at the 7 cm threshold.", out.width='75%'}
knitr::include_graphics(here::here("figures", "combined_richness.png"))
```

@@

```{r nvTrend710BAlightTemp, echo=FALSE, fig.cap="Mean regeneration values across the 200 samples per site with the y-axis scaled differently by model, for the 200 sites against gradients of (a) total basal area; (b) climatic water balance; (c) seasonal degree-day sum. The values were split into ten bins; the red lines represent a GAM model showing the trend in the observed data. For comparing the magnitude of recruitment across models the same graphs with equally scaled y-axis can be found in Figure S7.", fig.show = 'hold', out.width="49%",fig.ncol = 2, fig.subcap=c('', '', ''), fig.show='hold', fig.align='left'}

@@

The EuFoRIa data [@käber2023] are exceptional, particularly with respect
to the number of records (number of sites and repeated measurements),
which is essential for capturing a highly 'noisy' process such as tree
regeneration. The use of this dataset for model benchmarking provided
novel insights on the ability of state-of-the-art models to accurately
simulate regeneration levels, species composition, and mortality in an
early stage of tree life, i.e., between a tree diameter of 7 and 10 cm.
Overall, by adopting this approach a much broader understanding resulted
than if we had used these data for model calibration: it is primarily
from the shortcomings of the models that we can gain ecological insights
[cf. @trugman2022].

### 2023-04-28 manuscript/supplementaryRW.Rmd
added more changes from Haralds feedback

We simulated forest dynamics from bare ground for each site, using
patches (pixels) with a size of 20x20 cm^2^. We prepared landscapes of
250x250 m without topographical variation for each site. The simulations
we run from bare ground to 2000 years, and the samples were taken
starting in the year 2000. In the regeneration module, we used (1) a
base seeding probability of 0.001, which defines if seeds are available
due to background seed rain; (2) a dispersal mode based on the
formulation \'random asymmetric\'; and (3) the same establishment
probability of 0.9 for all species included in the simulation of, which
sets the same probability for all the species to regenerate
successfully, provided that the abiotic and light conditions are
suitable for establishment.

>>>>

We simulated forest dynamics from bare ground for each site, using
patches (pixels) with a size of 20x20 cm^2^. We prepared landscapes of
250x250 m without topographical variation for each site. The simulations
we run from bare ground to 2000 years, and the samples were taken
starting in the year 2000. In the regeneration module, we used (1) a
base seeding probability of 0.001, which defines if seeds are available
due to background seed rain; (2) a dispersal mode based on the
formulation 'random asymmetric'; and (3) the same establishment
probability of 0.9 for all species included in the simulation of, which
sets the same probability for all the species to regenerate
successfully, provided that the abiotic and light conditions are
suitable for establishment.
